Aligning species range data to better serve science and conservation

### 2016-01-16 ms_1_abst_sig.Rmd
a few little tweaks to wording, and incorporated a few more of Julie's comments

To understand mechanistic reasons for differences in species range predictions, we compare the distribution and extent of the pair of predicted ranges for the subset of species included in both datasets.  Categorizing the results based on how well each pair of species maps aligns, we identify several issues that commonly lead to misalignment in species range predictions.  Notably, IUCN maps often disregard bathymetry for depth-limited species, leading to predictions of species presence at unsuitable depths, and AquaMaps maps for data-poor species often extrapolate presence far afield from known occurrences.  Understanding these issues points toward solutions for providers and users of these datasets.

>>>>

For the subset of species included in both datasets, we compare the distribution and extent of the pair of predicted ranges and identify several issues that commonly lead to misalignment between the two datasets.  Notably, IUCN maps often disregard bathymetry for depth-limited species, leading to predictions of species presence at unsuitable depths, and AquaMaps maps for data-poor species often extrapolate presence far afield from known occurrences.  Understanding these issues points toward solutions for providers and users of these datasets.

### 2016-01-16 ms_2_intro_res_disc.Rmd
a few little tweaks to wording, and incorporated a few more of Julie's comments

@@

_Peer beneath the waves anywhere in the world - a moody kelp forest, a bustling coral reefscape, the frigid depths beneath a polar ice sheet, the endless blue of the open ocean - and you will find a unique ecosystem defined as much by which species are absent as by which species are present._ Knowing where individuals of a species exist, and where they thrive, is fundamental to the sciences of ecology, biogeography, and conservation, among many others.  This knowledge provides foundational information for 
 understanding species ranges and diversity,
 predicting species responses to human impacts and climate change,
 and managing and protecting species effectively.
A rich literature tackles the many dimensions of these questions.

@@

To understand the implications of differences between these two datasets, we compared how each data set represents the global spatial and taxonomic distribution of species.  For the relatively small number of species mapped in both datasets, we examined how well the species maps align and determined several issues that lead to misalignment between predicted species distributions. We then subjected two recent marine biodiversity studies - the Ocean Health Index biodiversity goal (Halpern et al. 2012) and a global analysis of gaps in protection afforded by marine protected areas (MPAs) (Klein et al. 2015) - to a sensitivity analysis, substituting one dataset over the other, to highlight the possible consequences of different data use decisions on our understanding of _the status of_ marine biodiversity.

@@

``` {r calc species numbers, echo = FALSE}
n_spp_am    <- nrow(spp_am)
n_spp_am_reviewed <- nrow(spp_am %>% filter(reviewed == 1))
p_spp_am_reviewed <- round(n_spp_am_reviewed/n_spp_am * 100, 1)
n_spp_am_rl <- nrow(spp_am %>% filter(sciname %in% spp_iucn_marine$sciname))
p_spp_am_rl <- round(n_spp_am_rl/n_spp_am * 100, 1)
n_spp_iucn  <- nrow(spp_iucn_maps)
n_spp_both  <- nrow(spp_map_pairs)
p_both_am   <- round(n_spp_both/n_spp_am * 100, 1)
p_both_iucn <- round(n_spp_both/n_spp_iucn * 100, 1)
```

The IUCN publishes species range maps as spatial vector polygon shapefiles, bundled by taxonomic groups.  Using GIS, experts outline spatial polygons to represent a given species' extent of occurrence, based on observation records and refined by expert understanding of the species' range and habitat preferences.  _IUCN releases a range map bundle for a taxonomic group once it has been "comprehensively assessed," i.e. at least 90% of the species within the taxonomic group have been evaluated (REF)._  While this mitigates sampling bias within taxa, it also means that entire taxonomic groups remain unavailable until they have met this threshold of comprehensive assessment.

@@

>>>>

``` {r calc species numbers, echo = FALSE}
n_spp_am    <- nrow(spp_am)
n_spp_am_reviewed <- nrow(spp_am %>% filter(reviewed == 1))
p_spp_am_reviewed <- round(n_spp_am_reviewed/n_spp_am * 100, 1)
n_spp_am_rl <- nrow(spp_am %>% filter(sciname %in% spp_iucn_marine$sciname))
p_spp_am_rl <- round(n_spp_am_rl/n_spp_am * 100, 1)
n_spp_iucn  <- nrow(spp_iucn_maps)
n_spp_both  <- nrow(spp_map_pairs)
p_both_am   <- round(n_spp_both/n_spp_am * 100, 1)
p_both_iucn <- round(n_spp_both/n_spp_iucn * 100, 1)
p_both_all  <- round(n_spp_both/nrow(spp_all) * 100, 1)
```

@@

_Peer beneath the waves anywhere in the world - a moody kelp forest, a bustling coral reefscape, the frigid depths beneath a polar ice sheet, the endless blue of the open ocean - and you will find a unique ecosystem defined as much by which species are absent as by which species are abundant._ Knowing where individuals of a species thrive is fundamental to the sciences of ecology, biogeography, and conservation, among many others.  This knowledge provides foundational information for 
 understanding species ranges and diversity,
 predicting species responses to human impacts and climate change,
 and managing and protecting species effectively.
A rich literature tackles the many dimensions of these questions.

@@

To understand the implications of differences between the AquaMaps and IUCN datasets, we compared how each data source represents the global spatial and taxonomic distribution of species.  For the `r p_both_all`% of species mapped in both datasets, we examined how well the species maps align and determined several issues that lead to misalignment between predicted species distributions, and outlined possible improvements.

We then subjected two recent marine biodiversity studies - the Ocean Health Index biodiversity goal (Halpern et al. 2012) and a global analysis of gaps in protection afforded by marine protected areas (MPAs) (Klein et al. 2015) - to a sensitivity analysis, substituting one dataset over the other, to highlight the possible consequences of different data use decisions on our understanding of _the status of_ marine biodiversity. ___We found XXX... very briefly outline results?___

@@

The IUCN publishes species range maps as polygon shapefiles, bundled by taxonomic groups (REF).  Using a geographic information system (GIS), experts outline spatial polygons to represent a given species' extent of occurrence, based on observation records and refined by expert understanding of the species' range and habitat preferences.  _IUCN releases a range map bundle for a taxonomic group once it has been "comprehensively assessed," i.e. at least 90% of the species within the taxonomic group have been evaluated (REF)._  While this mitigates sampling bias within taxa, it also means that entire taxonomic groups remain unavailable until they have met this threshold of comprehensive assessment.

@@

-----
# Implications

___From JSL:___ _I think you need a broader discussion of implications. How many times has IUCN/Aquamaps been cited? How long have they been around? What scales are the studies? Then you can muse about how results/interpretations would be different depending on which dataset they used, and reference Fig 3’s._

_Then, say for example, the OHI uses these datasets together (don’t need to get too much into OHI background I think, maybe they don’t even need to know about the biodiversity goal?), but say that to try to combine them they use a threshold at 40% since others use this threshold (but check out the SOM for analyses about thresholds)_
